---
title: "Particles of reported speech and inference"
author: "Samira Verhees"
output: html_document
bibliography: ./data/features/ref_bibs/evidentiality.bib
link-citations: true
---

---------------------------------------
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE)

# packages

library(tidyverse)
library(lingtypology)
library(DT)

# load data

vill <- read_tsv("data/villages_old.csv") # villages dataset
meta <- read_tsv("data/meta_old.csv") # language metadata and colors
particles <- read_tsv("data/features/tables/particles.csv") # feature dataset

# preparation of data

vill <- vill[complete.cases(vill$lat),] # remove villages for which we do not have coordinates (yet)

meta_core <- meta %>% # remove idioms not (yet) recognized as distinct
  filter(core == "yes")

vill_meta <- merge(vill, meta_core, by = "lang") # merge villages and coordinates with language metadata

# set for reportative particles

particlesrep_core <- particles %>%
  filter(`core:rep` == "yes")

particlesrep_vill <- merge(vill_meta, particlesrep_core, by = "lang") # merge villages, coordinates, and language metadata with feature information

particlesrep_vill$datapoint <- "extrapolated datapoint"

# set for quotative particles

particlesquot_core <- particles %>%
  filter(`core:quot` == "yes")

particlesquot_vill <- merge(vill_meta, particlesquot_core, by = "lang") # merge villages, coordinates, and language metadata with feature information

particlesquot_vill$datapoint <- "extrapolated datapoint"

```

## Maps

See [chapter](particles.html).

### Map 1. Quotative particles{.tabset .tabset-fade .tabset-pills}

#### Map: Extrapolated data

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width = 9.5}

map.feature(lang.gltc(particlesquot_vill$glottocode),
            latitude = particlesquot_vill$lat,
            longitude = particlesquot_vill$lon,
            features = particlesquot_vill$lang, # color feature = language
            color = particlesquot_vill$lang_color,
            stroke.features = particlesquot_vill$`value:quot`, # stroke.feature = your feature value
            stroke.color = c("black", "white"), 
            label = particlesquot_vill$lang,
            zoom.control = T,
            popup = paste("<b>Village:</b>", particlesquot_vill$village, "<br>", 
                          "<b>Source:</b>", particlesquot_vill$source, particlesquot_vill$page, "<br>",
                          "<b>Datapoint:</b>", particlesquot_vill$datapoint),
            width = 3, stroke.radius = 8,
            legend = FALSE)
```

#### Map: General datapoints

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width = 9.5}

particlesquot_core_meta <- merge(particlesquot_core, meta_core, by = "lang")

particlesquot_core_meta$datapoint <- "general datapoint"

map.feature(lang.gltc(particlesquot_core_meta$glottocode),
            features = particlesquot_core_meta$lang,
            color = particlesquot_core_meta$lang_color,
            legend = FALSE,
            stroke.features = particlesquot_core_meta$`value:quot`,
            stroke.color = c("black", "white"),
            label = particlesquot_core_meta$lang,
            zoom.control = T,
            popup = paste("<b>Source:</b>", particlesquot_core_meta$source, particlesquot_core_meta$page, "<br>",
                          "<b>Datapoint:</b>", particlesquot_core_meta$datapoint))

```

### Map 2. Quotative particles: structure{.tabset .tabset-fade .tabset-pills}

#### Map: Extrapolated data

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width = 9.5}

map.feature(lang.gltc(particlesquot_vill$glottocode),
            latitude = particlesquot_vill$lat,
            longitude = particlesquot_vill$lon,
            features = particlesquot_vill$lang, # color feature = language
            color = particlesquot_vill$lang_color,
            stroke.features = particlesquot_vill$`value:structure`, # stroke.feature = your feature value
            stroke.color = c("white", "black", "deeppink"), 
            label = particlesquot_vill$lang,
            zoom.control = T,
            popup = paste("<b>Village:</b>", particlesquot_vill$village, "<br>", 
                          "<b>Source:</b>", particlesquot_vill$source, particlesquot_vill$page, "<br>",
                          "<b>Datapoint:</b>", particlesquot_vill$datapoint),
            width = 3, stroke.radius = 8,
            legend = FALSE)
```

#### Map: General datapoints

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width = 9.5}

map.feature(lang.gltc(particlesquot_core_meta$glottocode),
            features = particlesquot_core_meta$lang,
            color = particlesquot_core_meta$lang_color,
            legend = FALSE,
            stroke.features = particlesquot_core_meta$`value:structure`,
            stroke.color = c("white", "black", "deeppink"),
            label = particlesquot_core_meta$lang,
            zoom.control = T,
            popup = paste("<b>Source:</b>", particlesquot_core_meta$source, particlesquot_core_meta$page, "<br>",
                          "<b>Datapoint:</b>", particlesquot_core_meta$datapoint))

```

### Map 3. Reportative particles{.tabset .tabset-fade .tabset-pills}

#### Map: Extrapolated data

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width = 9.5}

map.feature(lang.gltc(particlesrep_vill$glottocode),
            latitude = particlesrep_vill$lat,
            longitude = particlesrep_vill$lon,
            features = particlesrep_vill$lang, # color feature = language
            color = particlesrep_vill$lang_color,
            stroke.features = particlesrep_vill$`value:rep`, # stroke.feature = your feature value
            stroke.color = c("black", "white"), 
            label = particlesrep_vill$lang,
            zoom.control = T,
            popup = paste("<b>Village:</b>", particlesrep_vill$village, "<br>", 
                          "<b>Source:</b>", particlesrep_vill$source, particlesrep_vill$page, "<br>",
                          "<b>Datapoint:</b>", particlesrep_vill$datapoint),
            width = 3, stroke.radius = 8,
            legend = FALSE)
```

#### Map: General datapoints

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width = 9.5}

particlesrep_core_meta <- merge(particlesrep_core, meta_core, by = "lang")

particlesrep_core_meta$datapoint <- "general datapoint"

map.feature(lang.gltc(particlesrep_core_meta$glottocode),
            features = particlesrep_core_meta$lang,
            color = particlesrep_core_meta$lang_color,
            legend = FALSE,
            stroke.features = particlesrep_core_meta$`value:rep`,
            stroke.color = c("black", "white"),
            label = particlesrep_core_meta$lang,
            zoom.control = T,
            popup = paste("<b>Source:</b>", particlesrep_core_meta$source, particlesrep_core_meta$page, "<br>",
                          "<b>Datapoint:</b>", particlesrep_core_meta$datapoint))

```

### Map 4. Reportative particles: structure{.tabset .tabset-fade .tabset-pills}

#### Map: Extrapolated data

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width = 9.5}

map.feature(lang.gltc(particlesrep_vill$glottocode),
            latitude = particlesrep_vill$lat,
            longitude = particlesrep_vill$lon,
            features = particlesrep_vill$lang, # color feature = language
            color = particlesrep_vill$lang_color,
            stroke.features = particlesrep_vill$`value:structure`, # stroke.feature = your feature value
            stroke.color = c("white", "black", "deeppink"), 
            label = particlesrep_vill$lang,
            zoom.control = T,
            popup = paste("<b>Village:</b>", particlesrep_vill$village, "<br>", 
                          "<b>Source:</b>", particlesrep_vill$source, particlesrep_vill$page, "<br>",
                          "<b>Datapoint:</b>", particlesrep_vill$datapoint),
            width = 3, stroke.radius = 8,
            legend = FALSE)
```

#### Map: General datapoints

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width = 9.5}

map.feature(lang.gltc(particlesrep_core_meta$glottocode),
            features = particlesrep_core_meta$lang,
            color = particlesrep_core_meta$lang_color,
            legend = FALSE,
            stroke.features = particlesrep_core_meta$`value:structure`,
            stroke.color = c("white", "black", "deeppink"),
            label = particlesrep_core_meta$lang,
            zoom.control = T,
            popup = paste("<b>Source:</b>", particlesrep_core_meta$source, particlesrep_core_meta$page, "<br>",
                          "<b>Datapoint:</b>", particlesrep_core_meta$datapoint))

```

## Datatable

````{r}

# select which feature data you want to show in the datatable

dtable <- particles 

# generate searchable datatable

DT::datatable(dtable, escape = TRUE, rownames = TRUE,
              options = list(pageLength = 15, dom = 'ftip'))
```