---
title: "Perfect"
author: "Samira Verhees"
output: html_document
---

---------------------------------------

See [chapter](perfect.html).

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE)

library(tidyverse)
library(lingtypology)
library(RefManageR)

# make this part as a function and add it to example code

bib <- ReadBib(file = "./data/bibliography.bib")
BibOptions(bib.style = "authoryear", style = "html", cite.style = "authoryear", max.names = 2, sorting = "nyt")

# load data
fe <- read_csv("data/features/tables/perfects.csv") # feature dataset
mapname1 <- 'Functions of the perfect: resultative'
mapname2 <- 'Functions of the perfect: current relevance'
mapname3 <- 'Functions of the perfect: experiential'
mapname4 <- 'Functions of the perfect: indirect evidentiality'
mapname5 <- 'Structure of the perfect: synthetic vs. periphrastic'
mapname6 <- 'Structure of the perfect: lexical verb and auxiliary'
mapname7 <- 'Structure of the perfect: lexical verb and auxiliary'

```

# Map 1{.tabset .tabset-fade .tabset-pills}

`r mapname1`

## Map: Extrapolated data

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width = 9.5}
fe <- fe %>%
  mutate(value = value1)
source('./R/features_maker.R')
extr_map(fe)
```

## Map: General datapoints

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width = 9.5}
source('./R/features_maker.R')
gen_map(fe)
```

# Map 2{.tabset .tabset-fade .tabset-pills}

`r mapname2`

## Map: Extrapolated data

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width = 9.5}
fe <- fe %>%
  mutate(value = value2)
source('./R/features_maker.R')
extr_map(fe)
```

## Map: General datapoints

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width = 9.5}
source('./R/features_maker.R')
gen_map(fe)
```

# Map 3{.tabset .tabset-fade .tabset-pills}

`r mapname3`

## Map: Extrapolated data

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width = 9.5}
fe <- fe %>%
  mutate(value = value3)
source('./R/features_maker.R')
extr_map(fe)
```

## Map: General datapoints

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width = 9.5}
source('./R/features_maker.R')
gen_map(fe)
```

# Map 4{.tabset .tabset-fade .tabset-pills}

`r mapname4`

## Map: Extrapolated data

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width = 9.5}
fe <- fe %>%
  mutate(value = value4)
source('./R/features_maker.R')
extr_map(fe)
```

## Map: General datapoints

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width = 9.5}
source('./R/features_maker.R')
gen_map(fe)
```

# Map 5{.tabset .tabset-fade .tabset-pills}

`r mapname5`

## Map: Extrapolated data

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width = 9.5}
fe <- fe %>%
  mutate(value = value5)
source('./R/features_maker.R')
extr_map(fe)
```

## Map: General datapoints

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width = 9.5}
source('./R/features_maker.R')
gen_map(fe)
```

# Map 6{.tabset .tabset-fade .tabset-pills}

`r mapname6`

## Map: Extrapolated data

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width = 9.5}
fe <- fe %>%
  mutate(value = value6)
source('./R/features_maker.R')
extr_map(fe)
```

## Map: General datapoints

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width = 9.5}
source('./R/features_maker.R')
gen_map(fe)
```

# Map 7{.tabset .tabset-fade .tabset-pills}

`r mapname7`

## Map: Extrapolated data

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width = 9.5}
fe <- fe %>%
  mutate(value = value7)
source('./R/features_maker.R')
extr_map(fe)
```

## Map: General datapoints

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width = 9.5}
source('./R/features_maker.R')
gen_map(fe)
```

# Datatable

```{r, echo=FALSE, warning=FALSE, message=FALSE}
fe <- fe %>% 
  mutate(value1 = factor(value1)) %>% # change to value1
   mutate(value2 = factor(value2)) # change to value2
source('./R/features_maker.R')
d_b(fe, c('lang', 'family', 'aff', 'idiom', 'feature', 'value1', 'value2', 'value3', 'value4', 'value5', 'value6', 'value7', 'source', 'page'))
```

# References

```{r results = "asis", echo = FALSE}
PrintBibliography(bib, .opts = list(check.entries = FALSE))
```

<div class="tocify-extend-page" data-unique="tocify-extend-page" style="height: 0;"></div>
<!-- remove extra whitespace at bottom produced by floating table of contents and plots. 
Why do not work on GitHub?-->

