---
title: "The perfect"
author: "Samira Verhees"
output: html_document
bibliography: ./data/features/ref_bibs/evidentiality.bib
link-citations: true
---

---------------------------------------

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE)

# packages

library(tidyverse)
library(lingtypology)
library(leafsync)
library(DT)

# load data

vill <- read_tsv("data/villages_old.csv") # villages dataset
meta <- read_tsv("data/meta_old.csv") # language metadata and colors
perfect <- read_tsv("data/features/tables/perfects.csv") # feature dataset

# preparation of data

vill <- vill[complete.cases(vill$lat),] # remove villages for which we do not have coordinates (yet)

meta_core <- meta %>% # remove idioms not (yet) recognized as distinct
  filter(core == "yes")

vill_meta <- merge(vill, meta_core, by = "lang") # merge villages and coordinates with language metadata

perfect_core <- perfect %>%
  filter(core == "yes")

perfect_vill <- merge(vill_meta, perfect_core, by = "lang") # merge villages, coordinates, and language metadata with feature information

perfect_vill$datapoint <- "extrapolated datapoint"

```

## Maps

See the chapter [here](perfect.html).

### Map 1. Functions of the perfect: resultative{.tabset .tabset-fade .tabset-pills}

#### Map: Extrapolated data

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width = 9.5}

perfect_noresc <- perfect_vill[perfect_vill$resultative == "no",]

map.feature(lang.gltc(perfect_noresc$glottocode),
            latitude = perfect_noresc$lat,
            longitude = perfect_noresc$lon,
            features = perfect_noresc$lang, # color feature = language
            color = perfect_noresc$lang_color,
            stroke.features = perfect_noresc$`value:res`, # stroke.feature = your feature value
            stroke.color = c("black", "white"), 
            stroke.title = "Resultative",
            label = perfect_noresc$lang,
            zoom.control = T,
            popup = paste("<b>Village:</b>", perfect_noresc$village, "<br>", 
                          "<b>Source:</b>", perfect_noresc$source, perfect_noresc$page, "<br>",
                          "<b>Datapoint:</b>", perfect_noresc$datapoint),
            width = 3, stroke.radius = 8,
            legend = FALSE)

```

#### Map: General datapoints

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width = 9.5}

perfect_core_meta <- merge(perfect_core, meta_core, by = "lang")

perfect_core_meta$datapoint <- "general datapoint"

perfect_core_meta_noresc <- perfect_core_meta[perfect_core_meta$resultative == "no",]

map.feature(lang.gltc(perfect_core_meta_noresc$glottocode),
            features = perfect_core_meta_noresc$lang,
            color = perfect_core_meta_noresc$lang_color,
            legend = FALSE,
            stroke.features = perfect_core_meta_noresc$`value:res`,
            stroke.color = c("black", "white"),
            stroke.title = "Resultative",
            label = perfect_core_meta_noresc$lang,
            zoom.control = T,
            popup = paste("<b>Source:</b>", perfect_core_meta_noresc$source, perfect_core_meta_noresc$page, "<br>",
                          "<b>Datapoint:</b>", perfect_core_meta_noresc$datapoint))

```

### Map 2. Functions of the perfect: current relevance{.tabset .tabset-fade .tabset-pills}

#### Map: Extrapolated data

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width = 9.5}

map.feature(lang.gltc(perfect_noresc$glottocode),
            latitude = perfect_noresc$lat,
            longitude = perfect_noresc$lon,
            features = perfect_noresc$lang, # color feature = language
            color = perfect_noresc$lang_color,
            stroke.features = perfect_noresc$`value:cr`, # stroke.feature = your feature value
            stroke.color = c("black", "white"), 
            stroke.title = "Current relevance",
            label = perfect_noresc$lang,
            zoom.control = T,
            popup = paste("<b>Village:</b>", perfect_noresc$village, "<br>", 
                          "<b>Source:</b>", perfect_noresc$source, perfect_noresc$page, "<br>",
                          "<b>Datapoint:</b>", perfect_noresc$datapoint),
            width = 3, stroke.radius = 8,
            legend = FALSE)

```

#### Map: General datapoints

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width = 9.5}

map.feature(lang.gltc(perfect_core_meta_noresc$glottocode),
            features = perfect_core_meta_noresc$lang,
            color = perfect_core_meta_noresc$lang_color,
            legend = FALSE,
            stroke.features = perfect_core_meta_noresc$`value:cr`,
            stroke.color = c("black", "white"),
            stroke.title = "Current relevance",
            label = perfect_core_meta_noresc$lang,
            zoom.control = T,
            popup = paste("<b>Source:</b>", perfect_core_meta_noresc$source, perfect_core_meta_noresc$page, "<br>",
                          "<b>Datapoint:</b>", perfect_core_meta_noresc$datapoint))

```

### Map 3. Functions of the perfect: experiential{.tabset .tabset-fade .tabset-pills}

#### Map: Extrapolated data

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width = 9.5}

map.feature(lang.gltc(perfect_noresc$glottocode),
            latitude = perfect_noresc$lat,
            longitude = perfect_noresc$lon,
            features = perfect_noresc$lang, # color feature = language
            color = perfect_noresc$lang_color,
            stroke.features = perfect_noresc$`value:exp`, # stroke.feature = your feature value
            stroke.color = c("black", "white"), 
            stroke.title = "Experiential",
            label = perfect_noresc$lang,
            zoom.control = T,
            popup = paste("<b>Village:</b>", perfect_noresc$village, "<br>", 
                          "<b>Source:</b>", perfect_noresc$source, perfect_noresc$page, "<br>",
                          "<b>Datapoint:</b>", perfect_noresc$datapoint),
            width = 3, stroke.radius = 8,
            legend = FALSE)

```

#### Map: General datapoints

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width = 9.5}

map.feature(lang.gltc(perfect_core_meta_noresc$glottocode),
            features = perfect_core_meta_noresc$lang,
            color = perfect_core_meta_noresc$lang_color,
            legend = FALSE,
            stroke.features = perfect_core_meta_noresc$`value:exp`,
            stroke.color = c("black", "white"),
            stroke.title = "Experiential",
            label = perfect_core_meta_noresc$lang,
            zoom.control = T,
            popup = paste("<b>Source:</b>", perfect_core_meta_noresc$source, perfect_core_meta_noresc$page, "<br>",
                          "<b>Datapoint:</b>", perfect_core_meta_noresc$datapoint))

```

### Map 4. Functions of the perfect: indirect evidentiality{.tabset .tabset-fade .tabset-pills}

#### Map: Extrapolated data

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width = 9.5}

map.feature(lang.gltc(perfect_noresc$glottocode),
            latitude = perfect_noresc$lat,
            longitude = perfect_noresc$lon,
            features = perfect_noresc$lang, # color feature = language
            color = perfect_noresc$lang_color,
            stroke.features = perfect_noresc$`value:indir`, # stroke.feature = your feature value
            stroke.color = c("black", "white"), 
            stroke.title = "Indirect evidentiality",
            label = perfect_noresc$lang,
            zoom.control = T,
            popup = paste("<b>Village:</b>", perfect_noresc$village, "<br>", 
                          "<b>Source:</b>", perfect_noresc$source, perfect_noresc$page, "<br>",
                          "<b>Datapoint:</b>", perfect_noresc$datapoint),
            width = 3, stroke.radius = 8,
            legend = FALSE)

```

#### Map: General datapoints

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width = 9.5}

map.feature(lang.gltc(perfect_core_meta_noresc$glottocode),
            features = perfect_core_meta_noresc$lang,
            color = perfect_core_meta_noresc$lang_color,
            legend = FALSE,
            stroke.features = perfect_core_meta_noresc$`value:indir`,
            stroke.color = c("black", "white"),
            stroke.title = "Indirect evidentiality",
            label = perfect_core_meta_noresc$lang,
            zoom.control = T,
            popup = paste("<b>Source:</b>", perfect_core_meta_noresc$source, perfect_core_meta_noresc$page, "<br>",
                          "<b>Datapoint:</b>", perfect_core_meta_noresc$datapoint))

```

### Map 5. Structure of the perfect: synthetic vs. periphrastic{.tabset .tabset-fade .tabset-pills}

#### Map: Extrapolated data

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width = 9.5}
  # some problems here
# map.feature(lang.gltc(perfect_noresc$glottocode),
#             latitude = perfect_noresc$lat,
#             longitude = perfect_noresc$lon,
#             features = perfect_noresc$lang,
#             color = perfect_noresc$lang_color,
#             stroke.features = perfect_core_meta_noresc$`value:structure`,
#             stroke.color = c("deeppink", "black", "white"), 
#             stroke.title = "Synthetic structure",
#             label = perfect_noresc$lang,
#             zoom.control = T,
#             popup = paste("<b>Village:</b>", perfect_noresc$village, "<br>", 
#                           "<b>Source:</b>", perfect_noresc$source, perfect_noresc$page, "<br>",
#                           "<b>Datapoint:</b>", perfect_noresc$datapoint),
#             width = 3, stroke.radius = 8,
#             legend = FALSE)

```

#### Map: General datapoints

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width = 9.5}

map.feature(lang.gltc(perfect_core_meta_noresc$glottocode),
            features = perfect_core_meta_noresc$lang,
            color = perfect_core_meta_noresc$lang_color,
            legend = FALSE,
            stroke.features = perfect_core_meta_noresc$`value:structure`,
            stroke.color = c("deeppink", "black", "white"), 
            stroke.title = "Structure",
            label = perfect_core_meta_noresc$lang,
            zoom.control = T,
            popup = paste("<b>Source:</b>", perfect_core_meta_noresc$source, perfect_core_meta_noresc$page, "<br>",
                          "<b>Datapoint:</b>", perfect_core_meta_noresc$datapoint))

```

### Map 6. Structure of the perfect: lexical verb and auxiliary{.tabset .tabset-fade .tabset-pills}

#### Map: Extrapolated data

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width = 9.5}

lex <- map.feature(lang.gltc(perfect_noresc$glottocode),
            latitude = perfect_noresc$lat,
            longitude = perfect_noresc$lon,
            features = perfect_noresc$lang, # color feature = language
            color = perfect_noresc$lang_color,
            stroke.features = perfect_noresc$`value:lex`, # stroke.feature = your feature value
            stroke.color = c("black", "deeppink", "white"),
            stroke.title = "Lexical verb",
            label = perfect_noresc$lang,
            zoom.control = T,
            popup = paste("<b>Village:</b>", perfect_noresc$village, "<br>", 
                          "<b>Source:</b>", perfect_noresc$source, perfect_noresc$page, "<br>",
                          "<b>Datapoint:</b>", perfect_noresc$datapoint),
            width = 3, stroke.radius = 8,
            legend = FALSE)

aux <- map.feature(lang.gltc(perfect_noresc$glottocode),
            latitude = perfect_noresc$lat,
            longitude = perfect_noresc$lon,
            features = perfect_noresc$lang, # color feature = language
            color = perfect_noresc$lang_color,
            stroke.features = perfect_noresc$`value:aux`, # stroke.feature = your feature value
            stroke.color = c("black", "deeppink", "white"), 
            stroke.title = "Auxiliary verb",
            label = perfect_noresc$lang,
            zoom.control = T,
            popup = paste("<b>Village:</b>", perfect_noresc$village, "<br>", 
                          "<b>Source:</b>", perfect_noresc$source, perfect_noresc$page, "<br>",
                          "<b>Datapoint:</b>", perfect_noresc$datapoint),
            width = 3, stroke.radius = 8,
            legend = FALSE)

sync(lex, aux)

```

#### Map: General datapoints

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width = 9.5}

genlex <- map.feature(lang.gltc(perfect_core_meta_noresc$glottocode),
            features = perfect_core_meta_noresc$lang,
            color = perfect_core_meta_noresc$lang_color,
            stroke.features = perfect_core_meta_noresc$`value:lex`,
            stroke.color = c("black", "deeppink", "white"),
            stroke.title = "Lexical verb",
            label = perfect_core_meta_noresc$lang,
            zoom.control = T,
            popup = paste("<b>Village:</b>", perfect_core_meta_noresc$village, "<br>", 
                          "<b>Source:</b>", perfect_core_meta_noresc$source, perfect_core_meta_noresc$page, "<br>",
                          "<b>Datapoint:</b>", perfect_core_meta_noresc$datapoint),
            width = 3, stroke.radius = 8,
            legend = FALSE)

genaux <- map.feature(lang.gltc(perfect_core_meta_noresc$glottocode),
            features = perfect_core_meta_noresc$lang,
            color = perfect_core_meta_noresc$lang_color,
            stroke.features = perfect_core_meta_noresc$`value:aux`,
            stroke.color = c("black", "deeppink", "white"), 
            stroke.title = "Auxiliary verb",
            label = perfect_core_meta_noresc$lang,
            zoom.control = T,
            popup = paste("<b>Village:</b>", perfect_core_meta_noresc$village, "<br>", 
                          "<b>Source:</b>", perfect_core_meta_noresc$source, perfect_core_meta_noresc$page, "<br>",
                          "<b>Datapoint:</b>", perfect_core_meta_noresc$datapoint),
            width = 3, stroke.radius = 8,
            legend = FALSE)

sync(genlex, genaux)

```


## Datatable

````{r}

# select which feature data you want to show in the datatable

dtable <- perfect %>%
  select(lang, idiom, form, `value:lex`, `value:aux`, `value:structure`, `value:res`, `value:cr`, `value:exp`, `value:indir`, source, page)


# generate searchable datatable

DT::datatable(dtable, escape = TRUE, rownames = TRUE,
              options = list(pageLength = 15, dom = 'ftip'))
```